Preventing Runtime Errors of Redis at Compile Time

### 2015-11-19 paper.tex
Polish some of the wordings, up to Section 1.3.

@@

\title{Preventing runtime errors of Redis at compile time}

@@

\begin{abstract}
Programmers often interact with database systems, by sending queries through
 libraries or packages in some programming languages. But people make mistakes,
 while some of the syntactic and semantic errors can be prevented by the
 language at compile time, or caught by the package at runtime, most semantic
 errors are just being ignored, causing problems at the database system.

@@

Redis is an open source, in-memory data structure store, used as
 database, cache and message broker. Each value is associated with a binary-safe
 string key to identify and manipulate with. Redis supports many different
 kind of values, such as strings, hashes, lists, sets, etc.

To manipulate these values, Redis comes with a set of atomic operations, called
 \emph{commands}.

For example, if we want to add a bunch of strings to 2 sets, and intersect them,
 we could input the following sequence of commands into Redis's client
 interactively.

\begin{minted}[xleftmargin=1em]{text}
redis> SADD some-set a b c
(integer) 3
redis> SADD another-set a b
(integer) 2
redis> SINTER some-set another-set
1) "a"
2) "b"
\end{minted}

Keys such as \mintinline{text}{some-set} and \mintinline{text}{another-set}
 are created on site, with the command \mintinline{text}{SADD}, which returns
 the size of set after completed.

\subsection{Hedis}
Redis can also be used in most programming languages, with 3rd party libraries
 or packages that talk with Redis' TCP protocol. In Haskell, the most popular
 package is Hedis.

The previous example in Redis's client would look something like this with
 Hedis in Haskell.

@@

\mintinline{haskell}{sadd} takes a key and a list of values as arguments, and
 returns an \mintinline{haskell}{Integer}, wrapped in
 \mintinline{haskell}{Either} to indicate possible failures in the context
 \mintinline{haskell}{Redis}\footnotemark of command execution.

@@

Keys and values are of type \mintinline{haskell}{ByteString}, since they are all
 just binary strings in Redis. If a user wants to store values of
 arbitrary types, he or she will have to encode and decode them as
 \mintinline{haskell}{ByteString}s.

\footnotetext{Hedis provides another kind of context, \mintinline{haskell}{RedisTx}, for \emph{transactions}, united with \mintinline{haskell}{Redis} under the class of \mintinline{haskell}{RedisCtx}. We demonstrate \mintinline{haskell}{Redis} only for brevity. }

@@

Although everything in Redis is essentially a binary string, these strings are
 treated differently. Redis supports many different kind of data structures,
 such as strings, hashes, lists, etc. These values have different datatypes,
 and most commands only work with certain types of them, much like how C
 language treats a piece of data.

\paragraph{Problem 1} Consider the following case. \mintinline{text}{SET} is
 a command that only works with strings, it associates the key
 \mintinline{text}{some-string} with a value of string. If we treat the like a
 set and add an element to it with \mintinline{text}{SADD}, a runtime error
 arises.

@@

\paragraph{Problem 2} Even worse, not all strings are equal!
 \mintinline{text}{INCR} parses the string value as an integer, increments it by
 one, and store it back as a string. If a string value can't be parse as an
 integer, another runtime error arises

@@

\paragraph{In Hedis} The same goes for Hedis, since it's only a simple wrapping
 on top of Redis's protocol written in Haskell.

\begin{minted}[xleftmargin=1em]{haskell}
program :: Redis (Either Reply Integer)
program = do
    set "some-string" "foo"
    sadd "some-string" ["a"]
\end{minted}

It yields the same error as in Redis's client.

\begin{minted}[xleftmargin=1em]{haskell}
Left (Error "WRONGTYPE Operation against a
 key holding the wrong kind of value")
\end{minted}

>>>>

\newcommand{\todo}[2]{{\bf [#1]}: #2}

@@

\title{Preventing Runtime Errors of Redis at Compile Time}

@@

\begin{abstract}
Programmers often interact with database systems by sending queries through libraries or packages in some programming languages. People, however, make mistakes.
While some of the syntactic and semantic errors can be prevented by the
 language at compile time, or caught by the package at runtime, most semantic
 errors are just being ignored, causing problems at the database system.

@@

\todo{scm}{Cite some references for Redis?}

Redis is an open source, in-memory data structure store, often used as
database, cache and message broker. A Redis datatype can be think of as a set of key-value pairs, where each value is associated with a binary-safe string key to identify and manipulate with. Redis supports many different kind of values, such as strings, hashes, lists, and sets, etc, and provides a collection of of atomic \emph{commands} to manipulate these values.

For an example, consider the following sequence of commands, entered through the interactive interface of Redis. The keys
\mintinline{text}{some-set} abd \mintinline{text}{another-set} are
both associated to a set. The two call to command \mintinline{text}{SADD} respectively adds three and two values to the two sets, before \mintinline{text}{SINTER} takes their intersection:
\begin{minted}[xleftmargin=1em]{text}
redis> SADD some-set a b c
(integer) 3
redis> SADD another-set a b
(integer) 2
redis> SINTER some-set another-set
1) "a"
2) "b"
\end{minted}

Note that the keys \mintinline{text}{some-set} and \mintinline{text}{another-set}, if not existing before the call to \mintinline{text}{SADD}, are created on site. The call to
\mintinline{text}{SADD} returns the size of the set after
completion of the command.

\subsection{Hedis}

Many third party library exist to allow general purpose programmings
to access Redis databases through its TCP protocol. The most popular
such library for Haskell is Hedis.
%
\todo{scm}{Citation for Hedis?}
%
The following program is how the previous example looks like in
Hedis:

@@

The function \mintinline{haskell}{sadd} takes a key and a list of values as arguments, and
 returns an \mintinline{haskell}{Integer}, wrapped in
 \mintinline{haskell}{Either} to indicate possible failures in the context
 \mintinline{haskell}{Redis}\footnotemark of command execution.
\todo{scm}{Explain Reply in one sentence?}

@@

Note that keys and values, being nothing but binary strings in Redis,
are represented using Haskell \mintinline{haskell}{ByteString}.
Values of other types must be encoded as \mintinline{haskell}{ByteString}s before being written to the database,
and decoded after being read back.

\footnotetext{Hedis provides another kind of context, \mintinline{haskell}{RedisTx}, for \emph{transactions}, united with \mintinline{haskell}{Redis} under the class of \mintinline{haskell}{RedisCtx}. For brevity, we demonstrate only \mintinline{haskell}{Redis} in this paper. }

@@

\todo{scm}{Is this a quote? If so please cite the source. If not,
we shouldn't use the double quotes.}

Although everything in Redis is essentially a binary string, these strings are treated differently. Redis supports many different
kind of data structures, such as strings, hashes, lists, etc. While
they are all encoded as binary strings before being written to the databse, most commands, much like how the C language treats a piece of data, only works with data of certain types.

\paragraph{Problem 1} The command \mintinline{text}{SET}, by
its definition, associates a key to a string. In the following
example, the key \mintinline{text}{some-string} is associated
to string \mintinline{text}{foo}. Subsequent calls to
\mintinline{text}{SADD} causes runtime errors, since the value
of \mintinline{text}{some-string} is not a set, but a string.

@@

\paragraph{Problem 2} Even worse, not all strings are equal!
The call \mintinline{text}{INCR some-string} parses the string
associated with key \mintinline{text}{some-string}
to an integer, increments it by one, and store it back as a string.
If the string can not be parse as an integer, a runtime error
is raised.

@@

\paragraph{In Hedis} Hedis, being only a simple wrapper on top
of the TCP protocol of Redis, inherits all the problems mentioned
above. The following program yields the same error as that in
the Redis client.

\begin{minted}[xleftmargin=1em]{haskell}
program :: Redis (Either Reply Integer)
program = do
    set "some-string" "foo"
    sadd "some-string" ["a"]
\end{minted}
\begin{minted}[xleftmargin=1em]{haskell}
Left (Error "WRONGTYPE Operation against a
 key holding the wrong kind of value")
\end{minted}
